---
title: "Getting Started with autorelevate"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with autorelevate}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(autorelevate)
```

## Introduction

The autorelevate package implements the **autorelevated family** of
probability distributions, obtained by applying the autorelevation
(self-relevation) transform of Krakowski (1973) and Dileepkumar and
Sankaran (2022) to a baseline lifetime distribution. Ten baseline
families are supported: Weibull, Lomax, Burr XII, Gompertz,
Log-Logistic, Chen, Exponentiated Exponential, Power Lindley,
Log-normal, and Gamma. See `?autorelevate-package` for the transform's
derivation and `citation("autorelevate")` for full references.

## A Real Dataset

The package bundles `bladder_cancer`, the remission times (in months)
of 128 bladder cancer patients (Lee & Wang, 2003), the same dataset
analyzed by Dileep Kumar, Shabeer, and Sankaran (2025, Sec. 8.1).

```{r data}
data(bladder_cancer)
summary(bladder_cancer)
```

### Step 1: Diagnose the hazard shape with a TTT plot

Before fitting anything, a Total Time on Test (TTT) plot (Aarset, 1987)
shows whether the hazard rate is likely increasing, decreasing, bathtub,
or upside-down bathtub (UBT), purely from the shape of the empirical
curve relative to the 45-degree line.

```{r ttt}
ttt_plot(bladder_cancer)
```

The curve is concave then convex, indicating a UBT-shaped hazard --
exactly the shape the Autorelevated Weibull was designed to capture
(see `?haautorelevate` for the theorem governing when the ARW hazard is
increasing, decreasing, or UBT).

### Step 2: Compare all ten baseline families

```{r compare_families}
family_table <- compare_families(bladder_cancer)
print(family_table)
```

Families are ranked by AIC; `BIC`, `CAIC`, and `HQIC` are reported
alongside for cross-checking, since they penalize model complexity
differently (though here all ten families share the same two
parameters, so the ranking is driven purely by fit).

### Step 3: Fit and diagnose the best family

```{r fit, warning=FALSE, message=FALSE, fig.width=8, fig.height=4}
best_dist <- family_table$Family[1]
fit_best <- fit_autorelevate(bladder_cancer, dist = best_dist, method = "mle")
summary(fit_best)
plot(fit_best)
```

## Comparing Estimation Methods

Beyond comparing baseline families, you can compare MLE against Maximum
Product of Spacings (MPS), Least Squares (LS), Weighted Least Squares
(WLS), and Cramer-von Mises (CvM) for a single family:

```{r fit_all_methods}
fit_all_methods(bladder_cancer, dist = best_dist)
```

MPS is a robust alternative to MLE when the likelihood is unbounded or
the density has a singularity; LS and WLS (Swain, Venkatraman, & Wilson,
1988) are often preferred for small-to-moderate samples where MLE can be
unstable.

## Working With Distributions Directly

All ten baseline families expose the same six functions:
`dautorelevate()`, `pautorelevate()`, `sautorelevate()`,
`haautorelevate()`, `qautorelevate()`, and `rautorelevate()`.

```{r density}
x <- seq(0.01, 5, by = 0.05)
plot(x, dautorelevate(x, dist = "weibull", p1 = 0.5, p2 = 1.5), type = "l",
     ylab = "Density", main = "Autorelevated Weibull density")
```

```{r hazard}
plot(x, haautorelevate(x, dist = "weibull", p1 = 0.5, p2 = 0.8), type = "l",
     ylab = "Hazard rate", main = "Upside-down bathtub hazard (beta = 0.8)")
```

## References

Krakowski, M. (1973). The relevation transform and a generalization of
the gamma distribution function. *Revue francaise d'automatique,
informatique, recherche operationnelle. Recherche operationnelle*,
7(V2), 107-120.

Dileepkumar, M., & Sankaran, P. G. (2022). Some results of
auto-relevation transform in reliability analysis. *Statistics and
Applications*, 20(2), 251-263.

Dileep Kumar, M., Shabeer, A. M., & Sankaran, P. G. (2025). Reliability
properties and applications of autorelevated Weibull distribution.
*American Journal of Mathematical and Management Sciences*, 44(3-4),
215-237.

Aarset, M. V. (1987). How to identify a bathtub hazard rate. *IEEE
Transactions on Reliability*, R-36(1), 106-108.

Lee, E. T., & Wang, J. W. (2003). *Statistical Methods for Survival Data
Analysis* (3rd ed.). Wiley.

